---
title: "Covid19 vaccine study Fall2021"
author: "Keren Gonzalez"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 5
---


```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
TheData <-read.csv(file="FallFinalData.csv",header=TRUE) %>% select(-"X.1", -"X")
Covid91 <- TheData %>% mutate(across(where(is.character),as.factor)) 
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Part of the information you need to compute is the so-called effectiveness of the drug for each subgroup of people. Include a 95% confidence interval of the effectiveness for each different subgroup. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males

```{r}
Males <- filter(Covid91, sex == "Man")
```


```{r}
barchartGC(~infected + treatment,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")
```

The bar chart shows us that those who took the placebo got sick more than the males who took the Covid91. We can see that the pink bar for placebo is almost double in size compared to the drug. This indicates the drug is effective for the Males group.

## Numerical Summaries  
```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```

33.11% of the sick Males had the treatment, and 66.89% of those Males who got sick took the placebo. This is a 2 to 1 difference and lets us know that the treatment is in fact working.

## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Malesmod <- fisher.test(table1)
```

According to the Chi-squared test, the p-value is very small(2.2e-16). This indicates that we need to reject the null hypothesis and that there is a difference between the males who took the placebo and Males who took the drug.

```{r}
MaleEstimate <- Malesmod$estimate 
Malelow <- Malesmod$conf.int[1]
Malehigh <- Malesmod$conf.int[2]

Malelowprob <- Malelow / (1 + Malelow)
Maleloweffect <- Malelowprob * 100

Malehighprob <- Malehigh / (1+ Malehigh)
Malehigheffect <- Malehighprob * 100

Maleprob <- MaleEstimate / (1+ MaleEstimate)
MaleEffect <- Maleprob * 100
options(digits = 1)
```

The effectiveness of the Covid91 vaccine for Males is `r MaleEffect`. The confidence interval for males is `r Maleloweffect` to `r Malehigheffect`


# Females

```{r}
Females <- filter(Covid91, sex == "Woman")
```
```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")
```

According to the bar chart, this vaccine is not effective for Females. The blue bar that indicates that out of the Females who are sick more than 50% had taken the vaccine.

## Numerical Summaries  
```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```
This table reassures that the Females who got sick, 58% of them had taken the vaccine, versus 42% of the sick that took the placebo. Based on this information we can see that the vaccine seems to not be effective for Females.

## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Femalesmod <- fisher.test(table1)
```

Since our p-value is very small, we have to reject the null hypothesis, indicating that there is a difference between our data. In this case, it looks like the data is different in a bad way, as in the drug is not effective compared to the placebo.

## Calculating Effectiveness

```{r}
FemaleEstimate <- Femalesmod$estimate 
Femalelow <- Femalesmod$conf.int[1]
Femalehigh <- Femalesmod$conf.int[2]

Femalelowprob <- Femalelow / (1 + Femalelow)
Femaleloweffect <- Femalelowprob * 100

Femalehighprob <- Femalehigh / (1+ Femalehigh)
Femalehigheffect <- Femalehighprob * 100

Femaleprob <- FemaleEstimate / (1+ FemaleEstimate)
FemaleEffect <- Femaleprob * 100
options(digits = 1)
```

The effectiveness of the Covid91 vaccine for females is `r FemaleEffect`. The confidence interval for females is `r Femaleloweffect` to `r Femalehigheffect` . 


# Drug Users

```{r}
DrugUsers <- filter(Covid91, DrugUser == "yes")
```
```{r}
barchartGC(~infected + treatment,data=DrugUsers)
barchartGC(~infected + treatment,data=DrugUsers, type="percent")
```

# Numerical Summaries

```{r}
table1 <- xtabs(~infected + treatment, data=DrugUsers)
rowPerc(table1)
colPerc(table1)
```
## Inferential 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
DrugUsersmod <- fisher.test(table1)
```


## Calculating Effectiveness

```{r}
DrugUsersEstimate <- DrugUsersmod$estimate 
Druglow <- DrugUsersmod$conf.int[1]
Drughigh <- DrugUsersmod$conf.int[2]

Druglowprob <- Druglow / (1 + Druglow)
Drugloweffect <- Druglowprob * 100

Drughighprob <- Drughigh / (1+ Drughigh)
Drughigheffect <- Drughighprob * 100

Drugprob <- DrugUsersEstimate / (1+ DrugUsersEstimate)
DrugUserEffect <- Drugprob * 100
options(digits = 1)
```

The effectiveness of the Covid91 vaccine for drug users is `r DrugUserEffect`. The confidence interval for drug users is `r Drugloweffect` to `r Drughigheffect` . 

# LGBTQ

# Conclusions and recommendations

